---
title: "R Assignment 2"
author: "Jeff Moise"
date: "9/16/2019"
output:
  pdf_document: default
  html_document: default
---
Questions

1. Download the c2015 dataset to your computer.  Use function `getwd()` to check the current working directory. Use `setwd()` to change the current directory to the c2015 file. 
```{r}
work_dir <- "C://Users//student//Documents//RStudio"
getwd()
setwd(work_dir)
```


2. We need to install a package to read the xlsx file. (Let's not change the xlsx to csv here) There are a few packages for this.  I recommend to use the `readxl` package.  This package is contained in the `tidyverse` package so if you already installed `tidyverse`, you should have it already. If not, install and load the `readxl` package by

```{r, eval= FALSE}
install.packages('readxl') # install the library
library(readxl) # load the library
```

3.  Use `read_excel()` to read the c2015 dataset. Use function `class()` to check the type of data you just read in. You will notice that the data now is not just a data frame, it is also a `tibble`. A `tibble` is a generalization of a data frame, so you can still use all the functions and syntax for data frame with `tibble`. 
```{r}
path <- "C:/Users/student/Documents/RStudio/c2015.xlsx"
library(readxl)
data_excel=read_excel(path)
class(data_excel)
```


4. Use `dim` function to check the dimension of the data. Since this data is quite big, a common practice is to randomly subset the data to analyze.  Use `sample` function to create a new dataset that has a random 1000 observations from the original data. Use `set.seed(2019)` before using the `sample` function to set the seed for the randomness so that everyone in class is working with the same random subset of the data.  
```{r}

dim(data_excel)
set.seed(2019)
c2015sample<-data_excel[sample(1:80587,1000),]
dim(c2015sample)

```


5. Use `summary` function to have a quick look at the data. You will notice there is one variable is actually a constant.  Remove that variable from the data. 
```{r}
summary(c2015sample)
data_excel2 = subset(c2015sample, select = -c(YEAR) )
summary(data_excel2)
```

6. Check the number of missing values (NA) in each column. 
```{r}
colSums(is.na(data_excel2))
```


7. There are missing values in this data that are not NAs.  Identify the form of these missing values. Check the number of these missing values in each column.  Notice that you may want to use `na.rm = TRUE` when counting these missing values. 

```{r}
table(data_excel2$TRAV_SP)
table(data_excel2$OWNER)
table(data_excel2$DEFORMED)
```

8. Change the missing values in `SEX` variable to "Female"
```{r}
data_excel2$SEX[data_excel2$SEX=='Not Rep']= 'Female'
data_excel2$SEX[data_excel2$SEX=='Unknown']= 'Female'
table(data_excel2$SEX)
```


9. Fix the `AGE` variable so that it is in the right form and has no missing values.  **Hint:**
    - Change the value `Less than 1` to `0` (string 0, not a number 0)
    - Change the type of the variable to numeric using `as.numeric` function
    - Change the missing values to the average of the age. 
```{r}
data_excel2$SEX[data_excel2$AGE=='Less than 1']= '0'

data_excel2$AGE <- as.numeric(data_excel2$AGE)

avgAge<-mean(data_excel2$AGE[!is.na(data_excel2$AGE)])

data_excel2$AGE[is.na(data_excel2$AGE)]<- avgAge

avgAge




```
  
    
10. Put the `TRAV_SP`(Travel Speed) variable in the right form (type) and remove all missing values.  Calculate the average speed. You can use a non-base R function for this question. **Hint**: check out the function `str_replace`
```{r}
library(stringr)
data_excel2$TRAV_SP[data_excel2$TRAV_SP=='Stopped'] <- '0'
data_excel2$TRAV_SP[data_excel2$TRAV_SP=='Not Rep' | data_excel2$TRAV_SP=='Unknown'] <- NA
data_excel2$TRAV_SP<- stringr::str_replace(data_excel2$TRAV_SP," MPH", "")
data_excel2$TRAV_SP <- as.numeric(data_excel2$TRAV_SP)
mean(data_excel2$TRAV_SP, na.rm = TRUE)
```


11. Compare the average speed of those who had `"No Apprent Injury"` and the rest.  What do you observe? 
```{r}
data_excel2<- subset(data_excel2,data_excel2$INJ_SEV=='No Apparent Injury (O)') 
mean(data_excel2$TRAV_SP, na.rm=TRUE)



```


12. Use the `SEAT_POS` variable to filter the data so that there is only **drivers** in the dataset. Compare the average speed of man drivers and woman drivers. Comment on the results. 
```{r} 
table(data_excel2$SEAT_POS) 
data_excel2 <-subset(data_excel2,data_excel2$SEAT_POS=='Front Seat, Left Side') 
table(data_excel2$SEAT_POS)

MaleSet<-(subset(data_excel2,data_excel2$SEX=='Male'))
FemaleSet<-(subset(data_excel2,data_excel2$SEX=='Female'))
mean(MaleSet$TRAV_SP,na.rm = TRUE)
mean(FemaleSet$TRAV_SP,na.rm = TRUE)

# Males tend to drive faster slightly in this data set
                  

```


13. Compare the average speed of drivers who drink and those who do not. Comment on the results. **Hint:** This calculation can be done manually or by using the `aggregate` function  or `by` function in base R.  For example: 
```{r}
aggregate(data_excel2$TRAV_SP, by=list(data_excel2$DRINKING),FUN=mean, na.rm=TRUE)
```

14. Hypothesize about the age range of drivers who may drive more aggressively.  Test your hypothesis by comparing the average speed of those in this age range and the rest. Comment on the results.
```{r}
# I hypothesize that drivers under 30 would drive faster
data_excel2$AGE[data_excel2$AGE=='Unknown'] <- NA
over30 = subset(data_excel2$TRAV_SP,data_excel2$AGE>30)
under30 = subset(data_excel2$TRAV_SP,data_excel2$AGE<=30) 
over30 <- as.numeric(over30)
under30 <- as.numeric(under30) 
mean(over30,na.rm=TRUE)
mean(under30,na.rm=TRUE)

```


15. If the data did not confirm your hypothesis in 14.  Could you identify an age group of drivers who may drive more aggressively?  
```{r}
# I was unable to find a hige difference in driving speeds, it seems to random across the ages
aggregate(data_excel2$TRAV_SP, by=list(data_excel2$AGE),FUN=mean, na.rm=TRUE)
```


