---
title: "Assignment 1"
author: "Jeff Moise"
date: "9/9/2019"
output:
  pdf_document: default
  html_document: default
---

# 1. Calculate the Following Sum

S_{1} &= 1 + 2 +...+ 2019 \\ 
```{r}
s1= c(1:2019)
sum(s1)
```

S_{2} &= 1^3 + 2^3 +...+ 2019^3 \\
```{r}
s2= c(1:2019)
sum(s2^3)

```

S_{3} &= 1^1+2^2+3^3+...+2019^{2019} \\
```{r}
s3=c(1:2019)
sum(s3^s3)
```

S_{4} &= 1^1-2^2+3^3-4^4+...-2018^{2018}+2019^{2019} \\
```{r}
s4=c(1:2019)
b=s4^s4
c=c(1,-1)
sum(b * c)
```

S_{5} &= 1+frac{1}{4}+frac{1}{9}+frac{1}{16}+frac{1}{25}+... 
```{r}

s5=c(2:1000000)
sum(s5/(s5^2))
```

S_{6} &= 1+frac{1}{2}+frac{1}{3}+frac{1}{4}+...
```{r}
s6= c(1:1000000)
sum(1/s6)

```

S_{7} &= 1+frac{1}{8}+frac{1}{27}+frac{1}{64}+...
```{r}
s7 = c(1:100000)
sum(s7/s7^3)
```

S_{8} &=
1-frac{1}{2}+frac{1}{3}-frac{1}{4}+...
```{r}

s8=c(1:100000)
b=s8/(s8+1)
c=c(1,-1)
sum(b * c)

```
2. The `rnorm` function generate random variables from normal distribution. Generate a sample of 1000 values from normal distribution with the mean 10 and standard deviation 1.  
    a. Calculate the mean and standard deviation of the sample. 
    b. Out of 1000 samples, how many do you think are that great than 10?  Check your estimation.
    c. Use `hist()` function to show the histogram of the sample. 
    d  Estimate $P(X>1)$, where $X \sim N(2,1)$
```{r}
x=rnorm(1000,mean=10,sd=1)
mean(x)
sd(x)
b=x>10
sum(b)
hist(x)
y=rnorm(1000,mean=2,sd=1)
z=y>1
sum(z)/1000
```
3. Consider an experiment of tossing a fair dice.  
    a. Use the `sample` (with replacement) function to generate a sample of 1000 values from the experiment. 
    b. Calculate the mean and standard deviation of the sample.  
    c. How many times the 6 occured?
    d. Use `table` function to show the frequency of the values. 
    e. Use `prop.table(table())` to show the relative frequency of the values. 
    f. Plot the frequency of the values. 
```{r}
dice = c(1:6)
x=sample(dice,1000,replace=TRUE)
mean(x)
sd(x)
sum(x=6)
table(x)
prop.table(table(x)) 
plot(prop.table(table(x)))

```
4. Consider an experiment of tossing a dice 3 times. Let $X_1, X_2$, and $X_3$ be the number of tossing the first time, second time and third time, respectively. Use simulation to estimate the following probabilities:
    a. $P(X_1>X_2+X_3)$
    b. $P(X_1^2>X_2^2+X_3^2)$
```{r}
x=c(1:6)
x_1=sample(x,1000,replace=TRUE)
x_2=sample(x,1000,replace=TRUE)
x_3=sample(x,1000,replace=TRUE)
sum(x_1>x_2+x_3)/1000
sum(x_1^2>x_2^2+x_3^2)/1000

```
5. Using simulation, estimate the probability of getting three tails in a row when tossing a coin 3 times. **Hint:** one way is to generate a matrix with three columns where each rows is an observation of tossing a coin three times.
```{r}
x<-c(0:1)
a<-sample(x,1000,replace=TRUE)
b<-sample(x,1000,replace=TRUE)
c<-sample(x,1000,replace=TRUE)
z<-data.frame(a,b,c)
j<-rowSums(z)
sum(j==3)/1000




```
6. (**Extra Credits/Optional**) Using simulation, estimate the probability of getting three tails in a row when tossing a coin 10 times.



7.Central Limit Theorem (CLT). The CLT said that the mean of a sample of a distribution A (no matter what A is) follows normal distribution with the same mean as A. Following the below steps to confim the CLT when A is uniform distribution. 
    - Generate 100 samples of uniform distibution from 0 to 1.  Each sample has 1000 observations. Use the **runif** function to do this. 
    - Compute the means of the 100 samples. Create vector x containing these means. Hint:  You want to put all the samples in a matrix and use **rowSums** or **colSums** function. 
    - By CLT, x must follow normal distribution. Check this by plotting the histogram of x. Does it look like normal distribution? Use **hist(x)** to plot the histogram of x. 
    - Increase the number (100 and 1000) to see if the distribution of x looks more like normal distribution. 
    - Try the same procedure with two other distributions for A. 
```{r}

a<- matrix(runif(100*1000,0,1),ncol=100)

y <- rowSums(a)
means = y/1000

hist(means)

#Increase the numbers

a<-matrix(runif(1000*500,0,1),ncol=1000)

y<-rowSums(a)
means= y/1000
hist(means)

#two other distributions for A

a<-matrix(runif(100*1000,.7),ncol=1000)

y<-rowSums(a)
means= y/1000
hist(means)

a<-matrix(runif(100*1000,.2),ncol=1000)

y<-rowSums(a)
means= y/1000
hist(means)



```
    8. Use `read.csv` function to read in the titanic dataset.  You can find the dataset on Blackboard or at Kaggle.com.  Use `str` function to see a summary of the data.
  9. Use `knitr::kable` function to nicely print out the first 10 rows of the data in markdown.
```{r}
df <- read.csv(file="C:\\Users\\student\\Documents\\R\\titanic.csv")
str(df)
knitr::kable(head(df))
```
10. Use `is.na` function and `sum` function to count the total number of missing values in the data. Count the number of missing values in each columns. 
```{r}
sum(is.na(df))
```

10.Calculate the average Age of the passengers. You may want to use the parameter `na.rm = TRUE` in the function `mean`
```{r}
mean(df$Age, na.rm=TRUE)
```
11. Replace the missing values of age by the average age calculated previously. 
```{r}

df$Age.imp.mean <- ifelse(is.na(df$Age), mean(df$Age, na.rm=TRUE), df$Age)
sum(is.na(df$Age.imp.mean))
df$Age=df$Age.imp.mean
sum(is.na(df$Age))

```
12.Remove columns `Name`, `PassengerID`, `Ticket`, and `Cabin`. 
```{r}
drop <- c("Name","PassengerID","Ticket","Cabin")
df = df[,!(names(df) %in% drop)]
str(df)
```
13. Calculate the mean age of female passengers

```{r}
mean(df$Age[df$Sex=="female"])
```

14. Calculate the median fare of the passengers in Class 1
```{r}
median(df$Fare[df$Pclass=='1'])
```
15. Calculate the median fare of the female passengers that are not in Class 1
```{r}
median(df$Fare[df$Sex=='female' & df$Pclass!='1'])
```

16. Calculate the median age of survived passengers who are female and Class 1 or Class 2, 
```{r}
median(df$Age[df$Survived=='1' & df$Sex=='female' & (df$Pclass=='1' | df$Pclass=='2')])
```
17. Calculate the mean fare of female teenagers survived passengers
```{r}
mean(df$Fare[df$Survived=='1' & df$Sex=='female' & df$Age>12 &df$Age<20])
```
18.Calculate the mean fare of female teenagers survived passengers for each class
```{r}
df1 = df[((df$Survived==1)&(df$Sex=='female')& (df$Age>12) & (df$Age<20)),]

aggregate(df1$Fare,by=list(df1$Pclass),FUN=mean)


```
20.Calculate the ratio of Survived and not Survived for passengers who are who pays more than the average fare
```{r}
meanFare<-mean(df$Fare)
titanicsub= subset(df,Fare>meanFare)
sum(titanicsub$Survived ==1)/ sum(titanicsub$Survived==1 | titanicsub==0)
sum(titanicsub$Survived ==0)/ sum(titanicsub$Survived==1 | titanicsub==0)


```
21.	Add column that standardizes the fare (subtract the mean and divide by standard deviation) and name it `sfare`
```{r}
avgFare = mean(df$Fare)
sdFare = sd(df$Fare)
df$sfare <- (df$Fare-avgFare)/sdFare
head(df$sfare)
```
22.	Add categorical variable named `cfare` that takes value **cheap** for passengers paying less the average fare and takes value **expensive** for passengers paying more than the average fare. 
```{r}
df$cfare <- ifelse(df$Fare < avgFare, 'cheap','expensive')
head(df$cfare)
```
23. Add categorical variable named `cage` that takes value 0 for age 0-10, 1 for age 10-20, 2 for age 20-30, and so on
```{r}
df$cage <- 100
df$cage[df$Age<=10] <- 0
df$cage[df$Age>10 & df$Age<=20] <- 1
df$cage[df$Age>20 & df$Age<=30] <- 2
df$cage[df$Age>30] <- 3
head(df$cage)

```
24. Show the frequency of Ports of Embarkation. It appears that there are two missing values in the `Embarked` variable.  Assign the most frequent port to the missing ports. **Hint:** Use the `levels` function to modify the categories of categorical variables.   	


```{r}
summary(df$Embarked)
df$Embarked[df$Embarked==""] = "S"
summary(df$Embarked)
```

