---
title: "Assigment 5"
author: "Jeff Moise"
date: "9/30/2019"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This assignment works with the c2015 dataset.

1. Clean the data for easy graphing. Do the follows to clean and reduce the size of the data
Remove all observations that have a cell being either (1) NA, (2)‘Unknown’, (3)‘Not Rep’, or (4)‘Not Reported’
Remove all observations that have a cell containing either (1)‘Unknown’, (2)‘Not Rep’, or (3)‘Not Reported’. For instance, observations with DRINKING variable being Unknown (Police Reported) will be removed.
Fix TRAV_SP and AGE (following previous assignments) so that they are both numerics.
Filter so that there are only drivers in the data
```{r}

path <- "C:/Users/student/Documents/RStudio/c2015.xlsx"
library(readxl)
library(tidyverse)
library(dplyr)
library(stringr)
d=read_excel(path)
head(d)

#Remove NA, Unknown, Not Rep, Not Reported
d = d %>% filter_all(~!is.na(.))
d = d %>% filter_all(~!(.=="Unknown"))
d = d %>% filter_all(~!(.=="Not Rep"))
d = d %>% filter_all(~!(.==str_detect(.,"Not Rep")))
d = d %>% filter_all(~!(.==str_detect(.,"Unknown")))
d = d %>% filter_all(~!(.=="Not Reported"))
d = d %>% filter_all(~!(d$SEAT_POS == "Front Seat, Left Side"))

d$TRAV_SP[d$TRAV_SP=='Stopped'] <- '0'
d$TRAV_SP<- stringr::str_replace(d$TRAV_SP," MPH", "")
d$TRAV_SP <- as.numeric(d$TRAV_SP)


d<-d %>% 
  mutate(AGE=case_when(
    AGE=='Less than 1' ~ '0',
    TRUE ~ (AGE)))
d$AGE <- as.numeric(d$AGE)

head(d)


```


2. Use geom_point to plot AGE and TRAV_SP coloring by SEX.
```{r}
library(ggplot2)
ggplot(d, aes(AGE, TRAV_SP, color=SEX)) +
  geom_point()
```


3. There is overplotting in 2. Overplotting is when many points are duplicated on the graph. Use geom_jitter instead of geom_point for 2. to avoid overplotting.
```{r}
library(ggplot2)
ggplot(d, aes(AGE, TRAV_SP, color=SEX)) +
  geom_jitter()
```

4. Plot histograms of AGE, TRAV_SP with bins = 50.
```{r}
library(ggplot2)
ggplot(d, aes(AGE)) +
  geom_histogram(bins= 50)

ggplot(d, aes(TRAV_SP)) +
  geom_histogram( bins = 50)
```


5. Plot a histogram of AGE coloring (fill) by SEX.
```{r}

ggplot(d, aes(AGE, fill=SEX)) +
  geom_histogram()
```

6. Using geom_density to plot estimated densities of AGE colored by SEX.
```{r}
ggplot(d, aes(AGE, color=SEX)) +
  geom_density()
```

7. Plot estimated densities of TRAV_SP colored by INJ_SEV.
```{r}

ggplot(d, aes(TRAV_SP, color=INJ_SEV)) +
  geom_density()
```

8. Plot estimated densities of TRAV_SP seperated (colored) by weekdays and weekends.
```{r}
d<- d %>% 
  mutate(
   timeofweek = case_when( 
    DAY_WEEK=="Monday" ~ "weekday",
        DAY_WEEK=="Tuesday" ~ "weekday",
        DAY_WEEK=="Wednesday" ~ "weekday",
        DAY_WEEK=="Thursday" ~ "weekday",
        DAY_WEEK=="Friday" ~ "weekday",
    DAY_WEEK=="Saturday" ~ "weekend",
      DAY_WEEK=="Sunday" ~ "weekend"))

ggplot(d, aes(TRAV_SP, color=timeofweek)) +
  geom_density()

```

9. Implement geom_bar on MONTH. Implement geom_bar on MONTH filled by SEX
```{r}
ggplot(d, aes(MONTH, fill=SEX)) +
  geom_bar()
```


10. Implement geom_bar on MONTH and SEX with position='dodge'
```{r}
ggplot(d, aes(MONTH, fill=SEX)) +
  geom_bar(position="dodge")
```

11.
Plot a bar chart of average speeds in months using geom_col
```{r}
ggplot(d, aes(MONTH, TRAV_SP)) +
  stat_summary(fun.y = "mean", geom = "col")
```

12.
Plot a bar chart of average speeds in months using geom_bar
```{r}
ggplot(d, aes(MONTH, TRAV_SP)) +
  stat_summary(fun.y = "mean", geom = "bar")
```
13.
Plot a bar chart of average speeds in months filled by SEX
```{r}
ggplot(d, aes(MONTH, TRAV_SP, fill=SEX)) +
  stat_summary(fun.y = "mean", geom = "col")
```

14.
Plot a bar chart of average speeds in months colored by INJ_SEV
```{r}
ggplot(d, aes(MONTH, TRAV_SP, color=INJ_SEV)) +
  stat_summary(fun.y = "mean", geom = "col")
```


Refer to this link to have a similar following plot:

15.

Horizontal axis is for (monthly) average speed
The vertical axis is for months
Color by two colors: one for above overall average speed and the other for below the avarage speed
The speed on the horizontal axis is standardized

```{r}
library(ggplot2)

d1<- d %>% 
  group_by(MONTH) %>% 
  summarize(speed_avg = mean(TRAV_SP, na.rm=TRUE)) 
d1$speed_z= round((d1$speed_avg - mean(d1$speed_avg))/sd(d1$speed_avg),2)
d1$speed_type = ifelse(d1$speed_z<0, "below","above")
ggplot(d1,aes(x=reorder(MONTH, speed_z), y=speed_z, label=speed_z)) + 
  geom_bar(stat='identity', aes(fill=speed_type), width=.5) + scale_fill_manual(name="Speed",
          labels = c("Above Average", "Below Average"),
          values= c("above"="#00ba38", "below"="#f8766d")) + labs(subtitle= "Normalized Average Speeds by Month",
       title= "Diverging Bars") +
  coord_flip()
  

```

16. 
Refer to this link to have a similar following plot:
Horizontal Axis is for mean speed
Vertical Axis is for INJ_SEV
Color by SEX
The numbers of speed are shown in points.

```{r}

d1<- d %>% 
  group_by(INJ_SEV, SEX) %>% 
  summarize(speed_avg = mean(TRAV_SP, na.rm=TRUE))
d1$speed_z= round((d1$speed_avg - mean(d1$speed_avg))/sd(d1$speed_avg),2)
d1$speed_type = ifelse(d1$speed_z<0, "below","above")
ggplot(d1, aes(x=INJ_SEV, y=speed_z, label=speed_z, color=SEX))+
  geom_point(stat="identity", aes(col=speed_type), size=6) +
  scale_color_manual(name="Speed Avg",
          labels = c("Above Average", "Below Average"),
          values= c("above"="#00ba38", "below"="#f8766d"))+ geom_text(color="white", size=2) +
  labs(title="Diverging Dot Plot", subtitle= "Speed Avg by Injury Severity") + ylim(-2.5,2.5) + coord_flip()



```



17.
Refer to this link to have a similar following plot:
Horizontal Axis is for speed
Vertical Axis is for DAY
Color by SEX
The should be a invisible vertical line seperating the two sexes.

```{r}
df2= d %>% group_by(DAY,SEX) %>% summarize(SPEED=mean(TRAV_SP))
a=min(df2$SPEED)
b=max(df2$SPEED)

df2%>% mutate(SPEED1 = ifelse(SEX=='Female', (SPEED-a)/(a-b), (SPEED-a)/(b-a))) %>%
  ggplot(aes(x=reorder(DAY,-SPEED),y=SPEED1, fill=SEX)) + geom_col() + coord_flip()
```


18-20. Generate three other interesting graphs from the dataset.

```{r}
df2= d %>% group_by(STATE) %>% summarize(STATE_SPEED=mean(TRAV_SP)) %>% arrange(STATE_SPEED) %>%
  mutate(speed_cat=ifelse(STATE_SPEED>mean(d$TRAV_SP), 'Above Average', 'Below Average'))

df2 %>% mutate(STATE_SPEED = (STATE_SPEED - mean(STATE_SPEED))/sd(STATE_SPEED)) %>% 
  ggplot(aes(x=reorder(STATE,STATE_SPEED), y=STATE_SPEED, fill=speed_cat))+ 
  geom_col() +
  labs(x='DAYS') + coord_flip()
```
```{r}
d1<- d %>% 
  group_by(WEATHER) %>% 
  summarize(speed_avg = mean(TRAV_SP, na.rm=TRUE))
d1$speed_z= round((d1$speed_avg - mean(d1$speed_avg))/sd(d1$speed_avg),2)
d1$speed_type = ifelse(d1$speed_z<0, "below","above")
ggplot(d1, aes(x=WEATHER, y=speed_z, label=speed_z))+
  geom_point(stat="identity", aes(col=speed_type), size=6) +
  scale_color_manual(name="Speed Avg",
          labels = c("Above Average", "Below Average"),
          values= c("above"="#00ba38", "below"="#f8766d"))+ geom_text(color="white", size=2) +
  labs(title="Diverging Dot Plot", subtitle= "Speed Avg by Weather") + ylim(-2.5,2.5) + coord_flip()
```

```{r}
d1<- d %>% 
  group_by(DRINKING) %>% 
  summarize(speed_avg = mean(TRAV_SP, na.rm=TRUE))
d1$speed_z= round((d1$speed_avg - mean(d1$speed_avg))/sd(d1$speed_avg),2)
d1$speed_type = ifelse(d1$speed_z<0, "below","above")
ggplot(d1, aes(x=DRINKING, y=speed_z, label=speed_z))+
  geom_point(stat="identity", aes(col=speed_type), size=6) +
  scale_color_manual(name="Speed Avg",
          labels = c("Above Average", "Below Average"),
          values= c("above"="#00ba38", "below"="#f8766d"))+ geom_text(color="white", size=2) +
  labs(title="Diverging Dot Plot", subtitle= "Speed Avg Alcohol Involvment") + ylim(-2.5,2.5) + coord_flip()
```

