---
output: # rmarkdown::github_document
  pdf_document: default
  word_document: default
  html_document:
title: "Assignment 3.  Data Wrangling with Dplyr"
---
This assignment assumes that you have taken the `Introduction to the Tidyverse
` and `Data Manipulation with dplyr in R` course at Datacamp. You can use base R functions and dplyr functions in the assignment.

***Submission Instruction***.  You will need to submit on **Blackboard**, in the **Assignment** section, the follows:

- A knitted pdf
- A link to the markdown document in your Github
- A link to the pdf document in your Github

Questions
---

1. Read the `titanic` data set as a tibble. Redo questions 13 to 23 in the Assignment 1 using `dplyr`. **Notice:** you may want to use logical operators such as: 
```{r}
df <- read.csv(file="C:\\Users\\student\\Documents\\R\\titanic.csv") 
str(df)
```
13. Calculate the mean age of female passengers
```{r}

library("dplyr")          ## load
df %>%
  filter(Sex == "female") %>%
  summarize(mean(Age, na.rm = 1))
```
14. Calculate the median fare of the passengers in Class 1
```{r}
df %>%
  filter(df$Pclass== "1") %>%
  summarize(median(Fare, na.rm = 1))
```
Calculate the median fare of the female passengers that are not in Class 1
```{r}
df %>%
  filter(Sex == "female",Pclass!= "1" ) %>%
  summarize(median(Fare, na.rm = 1))
```
Calculate the median age of survived passengers who are female and Class 1 or Class 2, 
```{r}
df %>%
  filter(Sex == "female",Pclass== "1"|Pclass=="2",Survived=="1" ) %>%
  summarize(median(Age, na.rm = 1))
```
Calculate the mean fare of female teenagers survived passengers
```{r}
df %>%
  filter(Sex == "female", Age>12 & Age<20 ,Survived=="1" ) %>%
  summarize(mean(Fare, na.rm = 1))

```
Calculate the mean fare of female teenagers survived passengers for each class
```{r}
df %>%
  filter(Sex == "female", Age>12 & Age<20 ,Survived=="1" ) %>%
  group_by(Pclass) %>%
  summarize(mean(Fare, na.rm = 1))


```
Calculate the ratio of Survived and not Survived for passengers who are who pays more than the average fare
```{r}
meanFare<-mean(df$Fare)
titanicsub= subset(df,Fare>meanFare)

df %>% 
  filter(Fare>meanFare) %>% group_by(Survived) %>% summarise(n=n()) %>% mutate(f = n/sum(n))

```
Add column that standardizes the fare (subtract the mean and divide by standard deviation) and name it `sfare
```{r}
df = df %>% 
  mutate(sfare=Fare-mean(Fare)/sd(Fare))
names(df)
```
Add categorical variable named `cfare` that takes value **cheap** for passengers paying less the average fare and takes value **expensive** for passengers paying more than the average fare. 
```{r}
avgFare<- mean(df$Fare)

df = df %>% 
  mutate(cfare = case_when(
    Fare < avgFare ~ "cheap",
    Fare > avgFare ~ "expensive"))
names(df)
 
```
Add categorical variable named `cage` that takes value 0 for age 0-10, 1 for age 10-20, 2 for age 20-30, and so on
```{r}
df= mutate(
  df, 
  cage = case_when(
    Age <11 ~ "1",
    Age %in% 11:20 ~ "2",
     Age %in% 21:30 ~ "3",
     Age %in% 31:40 ~ "4",
     Age %in% 41:50 ~ "5",
     Age %in% 51:60 ~ "6",
    Age %in% 61:70 ~ "7",
    Age > 70 ~ "8"
    
  )
)
names(df)
```


```


|Operators|Discription|
|-------|--------------|
| !=    | not equal to |
| !x    | Not x        |
| x \text{|} y | x OR y|
| x & y | x AND y      |

2. Using Dplyr and in Assignment 2, redo  4 using `sample_n` function, redo 5 using `glimpse`, redo 11, 12 and 13. For 11, 12 and 13, you may want to use the combo `group_by` and `summarise`
 
 ----Use `dim` function to check the dimension of the data. Since this data is quite big, a common practice is to randomly subset the data to analyze.  Use `sample` function to create a new dataset that has a random 1000 observations from the original data. Use `set.seed(2019)` before using the `sample` function to set the seed for the randomness so that everyone in class 
 dim(data_excel)
set.seed(2019)
c2015sample<-data_excel[sample(1:80587,1000),]
dim(c2015sample) ---
```
```{r}
path <- "C:/Users/student/Documents/RStudio/c2015.xlsx"
library(readxl)
data_excel=read_excel(path)
class(data_excel)
```

```{r}
set.seed(2019)
df1<-sample_n(data_excel,1000, replace=TRUE)


```
Use `summary` function to have a quick look at the data. You will notice there is one variable is actually a constant.  Remove that variable from the data. 
summary(c2015sample)
data_excel2 = subset(c2015sample, select = -c(YEAR) )
summary(data_excel2)
```{r}
glimpse(df1)

df1<-select (df1,-c("YEAR"))
glimpse(df1)
```

Redo 11 from assignment 2
Compare the average speed of those who had `"No Apprent Injury"` and the rest.  What do you observe? 
data_excel2<- subset(data_excel2,data_excel2$INJ_SEV=='No Apparent Injury (O)') 
mean(data_excel2$TRAV_SP, na.rm=TRUE)
```{r}
library(stringr)
df1$TRAV_SP[df1$TRAV_SP=='Stopped'] <- '0'
df1$TRAV_SP[df1$TRAV_SP=='Not Rep' | df1$TRAV_SP=='Unknown'] <- NA
df1$TRAV_SP<- stringr::str_replace(df1$TRAV_SP," MPH", "")
df1$TRAV_SP <- as.numeric(df1$TRAV_SP)
avgSpeed<-mean(df1$TRAV_SP, na.rm = TRUE)

df1 %>% 
group_by(INJ_SEV) %>% 
  summarize(mean(TRAV_SP, na.rm = 1))
```
Redo 12 from assignment 2
Use the `SEAT_POS` variable to filter the data so that there is only **drivers** in the dataset. Compare the average speed of man drivers and woman drivers. Comment on the results. 

table(data_excel2$SEAT_POS) 
data_excel2 <-subset(data_excel2,data_excel2$SEAT_POS=='Front Seat, Left Side') 
table(data_excel2$SEAT_POS)

MaleSet<-(subset(data_excel2,data_excel2$SEX=='Male'))
FemaleSet<-(subset(data_excel2,data_excel2$SEX=='Female'))
mean(MaleSet$TRAV_SP,na.rm = TRUE)
mean(FemaleSet$TRAV_SP,na.rm = TRUE)
```{r}
library(dplyr)
df1 %>%
  filter(SEAT_POS=="Front Seat, Left Side") %>%
  group_by(SEX) %>%
  summarize(mean(TRAV_SP, na.rm=1))
```
redo 13 from assignment 2
 Compare the average speed of drivers who drink and those who do not. Comment on the results. **Hint:** This calculation can be done manually or by using the `aggregate` function  or `by` function in base R.  For example: 
 aggregate(data_excel2$TRAV_SP, by=list(data_excel2$DRINKING),FUN=mean, na.rm=TRUE)
```{r}
df1 %>%
  group_by(DRINKING) %>%
  summarize(mean(TRAV_SP, na.rm=1))
```

3. Calculate the travel speed (`TRAV_SP` variable) by day.  Compare the travel speed of the first 5 days and the last 5 days of months. 
```{r}
library(dplyr)

df1 %>% 
  mutate(
   timeofmonth = case_when( 
    DAY=="1" ~ "first 5",
        DAY=="2" ~ "first 5",
        DAY=="3" ~ "first 5",
        DAY=="4" ~ "first 5",
        DAY=="5" ~ "first 5",
    DAY=="27" ~ "last 5",
      DAY=="28" ~ "last 5",
      DAY=="29" ~ "last 5",
      DAY=="30" ~ "last 5",
      DAY=="31" ~ "last 5",)) %>%
  group_by(timeofmonth) %>%
  summarize(mean(TRAV_SP, na.rm=1))
```

4. Calculate the travel speed (`TRAV_SP` variable) by day of the week.  Compare the travel speed of the weekdays and weekends. 

```{r}

df1 %>% 
  mutate(
   timeofweek = case_when( 
    DAY_WEEK=="Monday" ~ "weekday",
        DAY_WEEK=="Tuesday" ~ "weekday",
        DAY_WEEK=="Wednesday" ~ "weekday",
        DAY_WEEK=="Thursday" ~ "weekday",
        DAY_WEEK=="Friday" ~ "weekday",
    DAY_WEEK=="Saturday" ~ "weekend",
      DAY_WEEK=="Sunday" ~ "weekend")) %>%
  group_by(timeofweek) %>%
  summarize(mean(TRAV_SP, na.rm=1))
```

5. Find the top 5 states with greatest travel speed.
```{r}
df1 %>%
  group_by(STATE) %>%
  summarize(mean(TRAV_SP, na.rm=1)) %>%
  top_n(5)
```


6. Rank the travel speed by `MONTH`.
```{r}

df1 %>%
  mutate(avgTrav=mean(TRAV_SP),rank= dense_rank(desc(avgTrav))) %>%
   group_by(rank, MONTH) %>%
               summarize(mean(TRAV_SP, na.rm=1))
 
              
                
```


7. Find the average speed of teenagers in December.
```{r}
df1 %>% 
  filter(MONTH=="December",AGE>12 & AGE<20) %>% 
  summarize(mean(TRAV_SP, na.rm=1))

```


8. Find the month that female drivers drive fastest on average.

```{r}
df1 %>% 
  filter(SEX=="Female") %>% 
  group_by(MONTH) %>%
  summarize(mean(TRAV_SP, na.rm=1)) %>%
  top_n(1)
```


9. Find the month that male driver drive slowest on average.
```{r}
df1 %>% 
  filter(SEX=="Male") %>% 
  group_by(MONTH) %>%
  summarize(mean(TRAV_SP, na.rm=1)) %>%
  top_n(-1)
```

10. Create a new column containing information about the season of the accidents. Compare the percentage of Fatal Injury by seasons. 
```{r}


unique(df1$INJ_SEV)
df1 %>% 
  mutate(
   season = case_when( 
      MONTH =="January" ~ "Winter",
       MONTH=="February" ~ "Winter",
       MONTH=="March" ~ "Spring",
        MONTH=="April" ~ "Spring",
        MONTH=="May" ~ "Spring",
        MONTH=="June" ~ "Summer",
        MONTH=="July" ~ "Summer",
        MONTH=="August" ~ "Summer",
        MONTH=="September" ~ "Fall",
        MONTH=="October" ~ "Fall",
       MONTH=="November" ~ "Fall",
       MONTH=="December" ~ "Winter")) %>% 
  group_by(season) %>%
  summarize(prop.table(table(INJ_SEV))[4])
```

11. Compare the percentage of fatal injuries for different type of deformations (`DEFORMED` variable)

```{r}
unique(df1$DEFORMED)
df1 %>% 
  group_by(DEFORMED) %>%
  summarize(prop.table(table(INJ_SEV))[4])

```



