---
output: # rmarkdown::github_document
  pdf_document: default
  word_document: default
  html_document:
title: "Assignment 4.  Data Wrangling with Dplyr- Jeff Moise"
---

Questions
---
```{r}
library(dplyr)
```

1. Compute the follows using `%>%` operator. **Notice that** 

  - `x %>%\ f = f(x)`
  - `x %>% f %>% g = g(f(x))` and 
  - `x %>% f(y) = f(x,y)`
  
  a. $\sin(2019)$
```{r}
2019 %>% sin()
```
  
  b. $\sin(\cos(2019))$ 
```{r}
2019 %>%
  sin() %>% cos()
```

  c. $\sin(\cos(\tan(\log(2019))))$ 

```{r}
2019 %>% sin() %>% cos() %>% tan() %>% log()
```

  d. $\log_2(2019)$
```{r}
2019 %>% log2()
```

2. Fixing the `SEX`, `AGE` and `TRAV_SP` following the steps in Assignment 2 (This time, do it on the entire dataset instead of the sample dataset). 
```{r}
path <- "C:/Users/student/Documents/RStudio/c2015.xlsx"
library(readxl)
df=read_excel(path)
class(df)

df<-df %>% 
  mutate(SEX=case_when(
    SEX=='Not Rep' ~ 'Female',
    SEX=='Unknown' ~ 'Female',
    SEX=='Male' ~ 'Male'))
table(df$SEX)

df<-df %>% 
  mutate(AGE=case_when(
    AGE=='Less than 1' ~ '0',
    TRUE ~ (AGE)))
df$AGE <- as.numeric(df$AGE)
table(df$AGE)

df$SEX[df$AGE=='Less than 1']= '0'
df$AGE <- as.numeric(df$AGE)
avgAge<-mean(df$AGE[!is.na(df$AGE)])
df$AGE[is.na(df$AGE)]<- avgAge
sum(is.na(df$AGE))


library(stringr)
df$TRAV_SP[df$TRAV_SP=='Stopped'] <- '0'
df$TRAV_SP[df$TRAV_SP=='Not Rep' | df$TRAV_SP=='Unknown'] <- NA
df$TRAV_SP<- stringr::str_replace(df$TRAV_SP," MPH", "")
df$TRAV_SP <- as.numeric(df$TRAV_SP)
mean(df$TRAV_SP, na.rm = TRUE)


```

3. Calculate the average age and average speed of female in the accident happened in the weekend. 
```{r}


df2<- df %>% 
  mutate(
   timeofweek = case_when( 
    DAY_WEEK=="Monday" ~ "weekday",
        DAY_WEEK=="Tuesday" ~ "weekday",
        DAY_WEEK=="Wednesday" ~ "weekday",
        DAY_WEEK=="Thursday" ~ "weekday",
        DAY_WEEK=="Friday" ~ "weekday",
    DAY_WEEK=="Saturday" ~ "weekend",
      DAY_WEEK=="Sunday" ~ "weekend"))

head(df2)

df2 %>% 
  filter(SEX=='Female' & DAY_WEEK=="weekend") %>% 
  summarize(mean(TRAV_SP, na.rm=TRUE))
```

_Notice: These questions are to practice `select_if` and `summarise_if`, `summarise_all`... functions in dplyr Check out the uses of these functions [here](https://dplyr.tidyverse.org/reference/select_all.html) and [here](https://dplyr.tidyverse.org/reference/summarise_all.html)._


4. Use `select_if` and `is.numeric` functions to create a dataset with only numeric variables. Print out the names of all numeric variables
```{r}
select_if(df2, is.numeric)
```

5. Calculate the mean of all numeric variables using `select_if` and `summarise_all` 
```{r}
df2 %>%
select_if(list(is.numeric)) %>%
summarise_all(mean)
```

6. We can shortcut 3 and 4 by using `summarise_if`:  Use `summarise_if` to Calculate the mean of all numeric variables. (You may need to use `na.rm = TRUE` to ignore the NAs)

```{r}
df2 %>%
summarise_if(is.numeric,mean, na.rm=TRUE)
```

7. Use `summarise_if` to calculate the median of all numeric variables.
```{r}
df2 %>% 
  summarise_if(is.numeric, median, na.rm=TRUE)
```


8. Use `summarise_if` to calculate the standard deviation of all numeric variables. (`sd` function for standard deviation)
```{r}
df2 %>% 
  summarise_if(is.numeric,sd,na.rm=TRUE)
```


9. Use `summarise_if` to calculate the number of missing values for each numeric variables. *Hint:* Use `~sum(is.na(.))` 
```{r}
df2 %>% 
  summarise_if(is.numeric, ~sum(is.na(.)))
```

10. Calculate the `log` of the average for each numeric variable. 
```{r}
df2 %>%
  summarise_if(is.numeric,~log(mean(.)))
```


11. You will notice that there is one NA is produced in Fix this by calculating the log of the absolute value average for each numeric variable. 

```{r}
df2 %>%
  summarise_if(is.numeric,~log(abs(mean(.))))
```


12. Calculate the number of missing values for each categorical variables using `summarise_if`
```{r}
df2 %>% 
  summarise_if(is.character, ~sum(is.na(.)))
```


13. Calculate the number of missing values for each categorical variables using `summarise_all`
```{r}
df2 %>%
select_if(is.character) %>%
summarise_all(~sum(is.na(.)))
```


14. Calculate the number of states in the dataset. **Hint: You can use `length(table())`
```{r}
df2 %>%
  summarise_all(~length(table(STATE)))

# could also use summarise_at()
df2 %>%
  summarise_at(c("STATE"),~length(table(STATE)))
```


15. Calculate the number of uniques values for each categorical variables using `summarise_if`.
```{r}
df2 %>% 
  summarise_if(is.character, ~length(table(.)))
```


16. Calculate the number of uniques values for each categorical variables using `summarise_all`.
```{r}
df2 %>% 
  select_if(is.character) %>%
  summarise_all(~length(table(.)))
```


17. Print out the names of all variables that have more than 30 distinct values
```{r}
names(df2 %>%
  select_if(~length(table(.))>30))
```


18. Print out the names of all categorical variables that more than 30 distinct values
```{r}
names(df2 %>% select_if(is.character) %>% 
       select_if(~length(table(.))>30))
```

19. Print out the names of all numeric variables that has the maximum values greater than 30
```{r}
names(df2 %>% select_if(is.numeric) %>% 
       select_if(~max(table(.))>30))
```

20. Calculate the mean of all numeric variables that has the maximum values greater than 30 using `summarise_if'
```{r}
df2 %>% select_if(is.numeric) %>% 
       summarize_if(~max(table(.))>30, mean, na.rm=1)
```

21. Calculate the mean of all numeric variables that has the maximum values greater than 30 using `summarise_all'
```{r}
names(df2 %>% select_if(is.numeric) %>%
        select_if(~max(table(.))>30) %>%
        summarize_all(mean,na.rm=1))
```


22. Create a dataset containing variables with standard deviation greater than 10. Call this data `d1`
```{r}
d0= names(df2 %>% select_if(is.numeric))[df2 %>% select_if(is.numeric) %>%
  summarise_all(sd, na.rm=1) > 10]

d1 = df2 %>% select(d0)

head(d1)
```


23. Centralizing a variable is subtract it by its mean. Centralize the variables of `d1` using `mutate_all`. Check the means of all centralized variables to confirm that they are all zeros. 
```{r}
d1 %>%
  mutate_all(funs(.-mean(.,na.rm=1))) %>%
  summarize_all(mean, na.rm=1)
```


24. Standarizing a variable is to subtract it to its mean and then divide by its standard deviation.  Standardize the variables of `d1` using `mutate_all`. Check the means and standard deviation of all centralized variables to confirm that they are all zeros (for the means) and ones (for standard deviation). 

```{r}
d1 %>%
  mutate_all(funs(.-mean(.,na.rm=1))) %>%
  mutate_all(funs(./sd(.))) %>%
  summarize_all(c(mean,sd), na.rm=1)
```



